March 23 Updates to report

### 2020-03-25 analysis/March23_updates.Rmd
some fixes for the technical report

```{r Figure 4b graph live BA over time}

@@

  # Plotting
  ggplot()+
  aes(x=Year,y=BA,color=Trt.Unit,group=Trt.Unit)+
  geom_point(size=3)+
  geom_line()+
  ylab("Mean basal area per plot (m²/ha)")+
  scale_x_continuous(limits=c(1992,2019),breaks=c(1992,1994,1997,2009,2019),
                     minor_breaks = c(1992,1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis
  facet_wrap(Trt.Unit~Plot,ncol=5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
        legend.position = "bottom")

@@

table4<-
  dat %>% 
   left_join(plotDat,by="Plot") %>% # join with plot data
  filter(Status=="Live") %>% 
  filter(Species!="Ep") %>% # remove birch from dataset
  mutate(Year=as.numeric(as.character(Year))) %>% 
  filter(Year %in% c(1994,2019)) %>% 
    mutate(BA=(DBH/100/2)^2*pi) %>% # convert tree DBH to basal area
 mutate(Trt.Unit=factor(paste("Trt:",BA.Target,sep=""))) %>% 
  group_by(Trt.Unit,Species,Year) %>% 
  summarise(BA=mean(BA,na.rm=TRUE)*20) %>% 
  mutate_at(vars("BA"),round,2) %>% 
  pivot_wider(names_from="Year",values_from="BA")

# Print to screen
  knitr::kable(table4,format="markdown")

@@

@@

  # Plotting
  ggplot()+
  aes(x=Year,y=BA.Mean,fill=Species)+
  geom_bar(stat="Identity")+
  facet_wrap(~label)+
  ylab("Mean basal area per plot (m2/ha)")+
  xlab("Measurement year")

@@

**NB**
However, this figure (and subsequent digging) has revealed another issue with the data - some plots were not measured in 1992, which that mean basal area contribution by species will vary between 1992 and 1994.  Note the decrease in Fd basal area per plot between 1992 and 1994.  This is due to different number of plots between 1992 and 1994, not because of Fd mortality (as this figure might imply).  I see a couple of options:

1. Explain this in the report; or
2. Limit analysis for these types of figures to 1994-2019 period.

@@

There is only one Fd tree in the entire dataset.

@@

sphCount<-
dat %>% 
    filter(Species!="Ep") %>% # remove birch 
  filter(Year %in% c(1992,1994)) %>% 
    filter(Status=="Live") %>% # filter for live only trees
   left_join(plotDat,by="Plot") %>% # join with plot data
   mutate(DBH.class=cut(DBH,diamClass,ordered_result = TRUE,right=F)) %>% # assign DBH classes
  group_by(BA.Target,Plot,Year,Species,DBH.class) %>% 
  summarise(Num.Trees=n()) %>% 
  ungroup() %>% 
  group_by(BA.Target,Year,Species,DBH.class) %>% 
    summarise(Num.Trees=mean(Num.Trees)) %>% 
  pivot_wider(names_from="Year",values_from = "Num.Trees") %>% 
arrange(BA.Target,Species,DBH.class)

@@

*This plot shows the mean number of stems per plot per Treatment Unit/Species/DBH Class/Year*

>>>>

```{r Figure 4b graph live BA over time with each PSP separate window}

@@

  # Plotting
  ggplot()+
  aes(x=Year,y=BA,color=Trt.Unit,group=Trt.Unit)+
  geom_point(size=3)+
  geom_line()+
  ylab("Total basal area per hectare (m²/ha)")+
  scale_x_continuous(limits=c(1992,2019),breaks=c(1992,1994,1997,2009,2019),
                     minor_breaks = c(1992,1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis
  facet_wrap(Trt.Unit~Plot,ncol=5)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
        legend.position = "bottom")

**Updated Figure 4b - All PSPs on one graph**

```{r Figure 4b graph live BA over time with each PSP on same graph}

  # Plotting
  ggplot()+
  aes(x=Year,y=BA,color=Trt.Unit,group=Plot)+
  geom_point(size=3)+
  geom_line()+
  ylab("Total basal area per hectare (m²/ha)")+
  scale_x_continuous(limits=c(1992,2019),breaks=c(1992,1994,1997,2009,2019),
                     minor_breaks = c(1992,1994,1997,2009,2019),
                     name="Measurement year")+ # stretch out x-axis
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5),
        legend.position = "bottom")

@@

# Create a matrix with number of plots (for calculating average BA)
table4.Numplots<-
    dat %>% 
   left_join(plotDat,by="Plot") %>% # join with plot data
  filter(Status=="Live") %>% 
  filter(Species!="Ep") %>% # remove birch from dataset
  mutate(Year=as.numeric(as.character(Year))) %>% 
  filter(Year %in% c(1994,2019)) %>% 
  mutate(Trt.Unit=factor(paste("Trt:",BA.Target,sep=""))) %>% 
  group_by(Trt.Unit,Year,Plot) %>% 
  summarise(BA=mean(DBH)) %>% 
  group_by(Trt.Unit,Year) %>% 
  summarise(num.plots=n())

table4<-
  dat %>% 
  left_join(plotDat,by="Plot") %>% # join with plot data
  filter(Status=="Live") %>% 
  filter(Species!="Ep") %>% # remove birch from dataset
  mutate(Year=as.numeric(as.character(Year))) %>% 
  filter(Year %in% c(1994,2019)) %>% 
  mutate(BA=(DBH/100/2)^2*pi) %>% # convert tree DBH to basal area
  mutate(Trt.Unit=factor(paste("Trt:",BA.Target,sep=""))) %>% 
  group_by(Trt.Unit,Species,Year) %>% 
  summarise(BA=sum(BA,na.rm=TRUE)) %>% # summarise BA for each treatment unit/year combination
  mutate(BA=BA*20) %>% # multiply by 20 to get per hectare

  # join with number of plots
  left_join(table4.Numplots,by=c("Trt.Unit","Year")) %>% 
  mutate(BA.ha=BA/num.plots) %>% # divide by number of plots to get mean BA by plot

  # remove columns for reshaping
  dplyr::select(-BA,-num.plots) %>%

  # pivot wider 
  pivot_wider(names_from="Year",values_from="BA.ha")

# Print to screen
  knitr::kable(table4,format="markdown",digits=2)

@@

Table shows mean BA per plot, multiplied by 20 to get BA/ha.

@@

  # Plotting
  ggplot()+
  aes(x=Year,y=BA.Mean,fill=Species)+
  geom_bar(stat="Identity")+
  facet_wrap(~label)+
  ylab("Mean basal area per plot (m²/ha)")+
  xlab("Measurement year")

@@

**NB**
Note the decrease in Fd basal area per plot between 1992 and 1994.  This is due to how I calculated these graphs (mean BA per plot), and the different number of plots between 1992 and 1994.  This is not because of Fd mortality (as this figure might imply).  I see a couple of options:

1. Explain this in the report; or
2. Limit analysis for these types of figures to 1994-2019 period; or
3. Come up with a different method to calculate BA by period (open to suggestions!).

@@

There is only one (albeit large) Fd tree in the entire dataset (note that there may be Fd in plots 16-19, which are not included in this analysis).

@@

sphCount<-
dat %>% 
    filter(Species!="Ep") %>% # remove birch 
  filter(Year %in% c(1992,1994)) %>% 
    filter(Status=="Live") %>% # filter for live only trees
   left_join(plotDat,by="Plot") %>% # join with plot data
   mutate(DBH.class=cut(DBH,diamClass,ordered_result = TRUE,right=F)) %>% # assign DBH classes
  group_by(BA.Target,Plot,Year,Species,DBH.class) %>% 
  summarise(Num.Trees=n()) %>% 
  ungroup() %>% 
  group_by(BA.Target,Year,Species,DBH.class) %>% 
    summarise(Num.Trees=mean(Num.Trees)*20) %>% # multiply by 20 to get per hectare
  pivot_wider(names_from="Year",values_from = "Num.Trees") %>% 
arrange(BA.Target,Species,DBH.class)

@@

*This plot shows the mean number of stems per hectare plot per Treatment Unit/Species/DBH Class/Year*
